% ============================================================
% abstract_en.tex — English Abstract
% ============================================================

\begin{abstracten}

As the context windows of large language models (LLMs) scale from
thousands to hundreds of thousands of tokens, KV-cache memory grows
linearly with sequence length and has become the primary bottleneck
for long-context inference. INT4 quantization can in principle
reduce KV-cache memory by a factor of four, yet existing methods
calibrated against numerical reconstruction error degrade sharply
in long-context settings: Needle retrieval accuracy drops from
100\% to 0\%, and perplexity (PPL) rises by more than 15\%. This
paper proposes the Kullback--Leibler divergence between pre- and
post-quantization attention distributions (\emph{attention-KL}) as
a unified principle for KV-cache quantization---defining the
calibration objective in the output behavior space rather than the
numerical-error space. With the grouped-query-attention (GQA)
key-value head count $H_{kv}$ as a structural organizing variable,
this single principle integrates calibration, diagnosis, and
fused-decoding efficiency.

We systematically validate the attention-KL principle on four main
models (Qwen2.5-1.5B/7B/14B and LLaMA-3.1-8B), with
Mistral-7B-Instruct-v0.3 as a third-family control, and summarize
the findings as three contributions.

\emph{First}, the attention-KL lens reveals a Key-dominated
degradation mechanism for INT4 KV-cache quantization. On
Qwen2.5-14B, a controlled ablation shows that retaining Key in
FP16 recovers 93\% of the PPL gap while retaining Value in FP16
recovers only 64\%; cross-model K-only/V-only ablations on four
main models consistently support this direction. The diagnosis
naturally derives an asymmetric format that combines per-channel-Key
and per-token-Value quantization---providing the first
attention-KL-based mechanistic account of the same format
independently proposed by KIVI~\cite{liu2024kivi} on engineering
grounds. At the calibration-objective level, we report a novel
"convergence--divergence--reconvergence" pattern between
attention-KL and mean squared error (MSE): under INT8 the two
objectives converge to the same $(p_c, g)$ point and produce
bit-identical calibration outputs (driven by the clip-rate $\leq 1\%$
robust-selection constraint); under INT4 they diverge on the 1.5B
model ($H_{kv}=2$) and re-converge on the 7B model ($H_{kv}=4$).

\emph{Second}, we instantiate the asymmetric format through an
offline attention-KL search and obtain \textbf{INT4-RoleAlign},
which replaces KIVI's runtime absmax/min statistics with a
reproducible, auditable offline calibration interface---producing
frozen quantization parameters suitable for compliance-sensitive
deployment. Across four main models, INT4-RoleAlign preserves
100\% Needle retrieval at a KV-cache compression ratio of roughly
73\%, with PPL degradation decreasing markedly with model scale
(1.5B: 13.7\%, 7B: 6.1\%, 8B: 2.4\%) and a small marginal
improvement over the same-format KIVI-style baseline on the
LongBench synthetic score (1.5B: +0.09, 7B: +0.06). External
validation on the 14B model further confirms a 100\% Needle pass
rate over 4K--32K contexts and RULER scores of 96.6--98.5\% over
4K--16K (14B PPL is measured on a 32K prefix and is reported as
an external-validity anchor rather than a cross-model comparison).

\emph{Third}, the Triton fused quantized-decoding kernel exhibits
a clear efficiency phase diagram across the $(H_{kv},
\text{seq\_len})$ space: at $H_{kv}=8$, a 32K context length
yields roughly 40\% decode-latency reduction relative to the
same-configuration INT4 reference implementation. The 8B-LLaMA
and 14B-Qwen models, sharing $H_{kv}=8$ across a 1.75$\times$
parameter gap and different model families, exhibit consistent
kernel-efficiency profiles---supporting $H_{kv}$ (rather than
model scale) as the structural driver of fused-kernel efficiency.
We convert this phase diagram into a layered deployment policy on
NVIDIA H20.

Together, the three contributions form an attention-KL-centered,
$H_{kv}$-structured behavior-aligned quantization framework
spanning diagnosis, calibration, and deployment.

\keywordsen{large language models; KV-cache quantization;
behavior-aligned calibration; asymmetric quantization;
grouped-query attention}

\end{abstracten}
